\chapter{Sentential Logic}
\markright{Chap. \ref{chap:SL}: Sentential Logic}
\label{chap:SL}
\setlength{\parindent}{1em}

This chapter introduces a logical language called SL. It is a version of \emph{sentential logic}, because the basic units of the language will represent statements, and a statement is usually given by a complete sentence in English.




% ******************************************
%  * Section 6.1  Sentence Letters                          *
% ******************************************

\section{Sentence Letters}


\newglossaryentry{sentence letter}
{
name=sentence letter,
description={A single capital letter, used in SL to represent a basic sentence.}
}

\newglossaryentry{symbolization key}
{
name=symbolization key,
description={A list that shows which English sentences are represented by which sentence letters in SL. It is also called a dictionary.}
}

In SL, capital letters, called \textsc{\glspl{sentence letter}} \label{def:sentence_letter} are used to represent simple statements. Considered only as a symbol of SL, the letter $A$ could mean any statement. So when translating from English into SL, it is important to provide a symbolization key, or dictionary. The \textsc{\gls{symbolization key}} \label{def:symbolization_key} provides an English language sentence for each sentence letter used in the symbolization.

Consider this argument:
\begin{earg}
\item[1.] There is an apple on the desk.
\item[2.] If there is an apple on the desk, then Jenny made it to class.
\item[] \textcolor{white}{.}\sout{\hspace{.8\linewidth}}\textcolor{white}{.} 
\item[$\therefore$] Jenny made it to class.
\end{earg}
This is obviously a valid argument in English. In symbolizing it, we want to preserve the structure of the argument that makes it valid.
What happens if we replace each sentence with a letter? Our symbolization key would look like this:
\begin{ekey}
\item[A:]There is an apple on the desk.
\item[B:]If there is an apple on the desk, then Jenny made it to class.
\item[C:]Jenny made it to class.
\end{ekey}
We would then symbolize the argument in this way:
\begin{earg}
\item[1.] $A$
\item[2.] $B$
\item[] \textcolor{white}{.}\sout{\hspace{.05\linewidth}}\textcolor{white}{.} 
\item[$\therefore$] $C$
\end{earg}
There is no necessary connection between some sentence $A$, which could be any statement, and some other sentences $B$ and $C$, which could also be anything.
The structure of the argument has been completely lost in this translation.

The important thing about the argument is that the second premise is not merely \emph{any} statement, logically divorced from the other statement in the argument. The second premise contains the first premise and the conclusion \emph{as parts}. Our symbolization key for the argument only needs to include meanings for $A$ and $C$, and we can build the second premise from those pieces. So we symbolize the argument this way:
\begin{earg}
\item[1.] $A$
\item[2.] If $A$, then $C$.
\item[] \textcolor{white}{.}\sout{\hspace{.2\linewidth}}\textcolor{white}{.} 
\item[$\therefore$] $C$
\end{earg}
This preserves the structure of the argument that makes it valid, but it still makes use of the English expression ``If$\ldots$ then$\ldots$.'' Although we ultimately want to replace all of the English expressions with logical notation, this is a good start.

\newglossaryentry{atomic sentence}
{
name=atomic sentence,
description={A sentence that does not have any sentences as proper parts.}
}


The individual sentence letters in SL are called atomic sentences, because they are the basic building blocks out of which more complex sentences can be built. We can identify atomic sentences in English as well. An \textsc{\gls{atomic sentence}} \label{def:atomic_sentence} is one that cannot be broken into parts that are themselves sentences. ``There is an apple on the desk'' is an atomic sentence in English, because you can't find any proper part of it that forms a complete sentence. For instance ``an apple on the desk'' is a noun phrase, not a complete sentence. Similarly ``on the desk'' is a prepositional phrase, and not a sentence, and ``is an'' is not any kind of phrase at all. This is what you will find no matter how you divide ``There is an apple on the desk.'' On the other hand you can find two proper parts of ``If there is an apple on the desk, then Jenny made it to class'' that are complete sentences: ``There is an apple on the desk'' and ``Jenny made it to class.'' As a general rule, we will want to use atomic sentences in SL (that is, the sentence letters) to represent atomic sentences in English. Otherwise, we will lose some of the logical structure of the English sentence, as we have just seen. 

%Atomic sentences go together to make complex sentences in much the same way that physical atoms go together to make molecules. Physical atoms were originally called `atoms' because chemists thought that they were irreducible. Chemists were wrong, and physical atoms can be split.

%It is important to remember that a symbolization key only gives the meaning of atomic sentences for purposes of translating a specific argument.

There are only 26 letters of the alphabet, but there is no logical limit to the number of atomic sentences. We can use the same letter to symbolize different atomic sentences by adding a subscript, a small number written after the letter. We could have a symbolization key that looks like this:
\begin{ekey}
\item[A$_1$:] The apple is under the armoire.
\item[A$_2$:] Arguments in SL always contain atomic sentences.
\item[A$_3$:] Adam Ant is taking an airplane from Anchorage to Albany.
\item[$\vdots$]
\item[A$_{294}$:] Alliteration angers otherwise affable astronauts.
\end{ekey}
Keep in mind that each of these is a different sentence letter. When there are subscripts in the symbolization key, it is important to keep track of them.


% ******************************************
%  * Sentential Connectives		                          *
% ******************************************

\section{Sentential Connectives}

\newglossaryentry{sentential connective}
{
name=sentential connective,
description={A logical operator in SL used to combine sentence letters into larger sentences.}
}

Logical connectives are used to build complex sentences from atomic components. In SL, our logical connectives are called \textsc{\glspl{sentential connective}} \label{def:sentential_connective} because they connect sentence letters. There are five sentential connectives in SL. This table summarizes them, and they are explained below.

\begin{table}[h]
\center
\begin{tabular}{|c|c|c|}
\hline
Symbol&What it is called&What it means\\
\hline
\enot&negation&``It is not the case that$\ldots$''\\
\eand&conjunction&``Both $\ldots$\ and $\ldots$''\\
\eor&disjunction&``Either $\ldots$\ or $\ldots$''\\
\eif&conditional&``If $\ldots$\ then $\ldots$''\\
\eiff&biconditional&``$\ldots$ if and only if $\ldots$''\\
\hline
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%% 2.2.1 Negation

\subsection{Negation}
Consider how we might symbolize these sentences:
\begin{earg}
\item[\ex{not1}] Mary is in Barcelona.
\item[\ex{not2}] Mary is not in Barcelona.
\item[\ex{not3}] Mary is somewhere other than Barcelona.
\end{earg}

In order to symbolize sentence \ref{not1}, we will need one sentence letter. We can provide a symbolization key:

\begin{ekey}
\item[B:]Mary is in Barcelona.
\end{ekey}

Note that here we are giving $B$ a different interpretation than we did in the previous section. The symbolization key only specifies what $B$ means \emph{in a specific context}. It is vital that we continue to use this meaning of $B$ so long as we are talking about Mary and Barcelona. Later, when we are symbolizing different sentences, we can write a new symbolization key and use $B$ to mean something else.

\newglossaryentry{negation}
{
name=negation,
description={The symbol \enot, used to represent words and phrases that function like the English word ``not''.}
}

Now, sentence \ref{not1} is simply $B$. Sentence \ref{not2} is obviously related to sentence \ref{not1}: it is basically \ref{not1} with a ``not'' added. We could put the sentence partly our symbolic language by writing ``Not $B$.'' This means we do not want to introduce a different sentence letter for \ref{not2}. We just need a new symbol for the ``not'' part. Let's use the symbol `\enot,' which we will call \textsc{\gls{negation}}. \label{def:negation} Now we can translate `Not $B$' to $\enot B$. 

Sentence \ref{not3} is about whether or not Mary is in Barcelona, but it does not contain the word ``not.'' Nevertheless, it is obviously logically equivalent to sentence \ref{not2}. They both mean: It is not the case that Mary is in Barcelona. As such, we can translate both sentence \ref{not2} and sentence \ref{not3} as $\enot B$.

\factoidbox{
A sentence can be symbolized as $\enot\script{A}$ if it can be paraphrased in English as ``It is not the case that \script{A}.''
}


Consider these further examples:
\begin{earg}
\item[\ex{not4}] The widget can be replaced if it breaks.
\item[\ex{not5}] The widget is irreplaceable.
\item[\ex{not5b}] The widget is not irreplaceable.
\end{earg}


If we let $R$ mean ``The widget is replaceable'', then sentence \ref{not4} can be translated as $R$.

What about sentence \ref{not5}? Saying the widget is irreplaceable means that it is not the case that the widget is replaceable. So even though sentence \ref{not5} is not negative in English, we symbolize it using negation as $\enot{R}$.

Sentence \ref{not5b} can be paraphrased as ``It is not the case that the widget is irreplaceable.'' Using negation twice, we translate this as $\enot\enot R$. The two negations in a row each work as negations, so the sentence means ``It is not the case that it is not the case that $R$.'' If you think about the sentence in English, it is logically equivalent to sentence \ref{not4}. So when we define logical equivalence in SL, we will make sure that $R$ and $\enot\enot R$ are logically equivalent.

More examples:
\begin{earg}
\item[\ex{not6}] Elliott is happy.
\item[\ex{not7}] Elliott is unhappy.
\end{earg}


If we let $H$ mean ``Elliot is happy'', then we can symbolize sentence \ref{not6} as $H$.

However, it would be a mistake to symbolize sentence \ref{not7} as $\enot{H}$. If Elliott is unhappy, then he is not happy---but sentence \ref{not7} does not mean the same thing as ``It is not the case that Elliott is happy.'' It could be that he is not happy but that he is not unhappy either. Perhaps he is somewhere between the two. In order to symbolize sentence \ref{not7}, we would need a new sentence letter.

For any sentence \script{A}: If \script{A} is true, then \enot\script{A} is false. If \enot\script{A} is true, then \script{A} is false. Using T for true and F for false, we can summarize this in a \emph{characteristic truth table} for negation:
\begin{center}
\begin{tabular}{c|c}
\script{A} & \enot\script{A}\\
\hline
T & F\\
F & T 
\end{tabular}
\end{center}
We will discuss truth tables at greater length in the next chapter.

%%%%%%%%%%%%%%%%%% 2.2.2 Conjunction

\subsection{Conjunction}
Consider these sentences:
\begin{earg}
\item[\ex{and1}]Adam is athletic.
\item[\ex{and2}]Barbara is athletic.
\item[\ex{and3}]Adam is athletic, and Barbara is also athletic.
\end{earg}

We will need separate sentence letters for \ref{and1} and \ref{and2}, so we define this symbolization key:
\begin{ekey}
\item[A:] Adam is athletic.
\item[B:] Barbara is athletic.
\end{ekey}

Sentence \ref{and1} can be symbolized as $A$.

Sentence \ref{and2} can be symbolized as $B$.

\newglossaryentry{conjunction}
{
name=conjunction,
description={The symbol \eand, used to represent words and phrases that function like the English word ``and.''}
}

\newglossaryentry{conjunct}
{
name=conjunct,
description={A sentences joined to another by a conjunction.}
}

Sentence \ref{and3} can be paraphrased as ``$A$ and $B$.'' In order to fully symbolize this sentence, we need another symbol. We will use \eand. We translate ``$A$ and $B$'' as $A\eand B$. The logical connective \eand is called the \textsc{\gls{conjunction}}, \label{def:conjunction} and $A$ and $B$ are each called \textsc{\glspl{conjunct}}. \label{def:conjunct}
 

Notice that we make no attempt to symbolize ``also'' in sentence \ref{and3}. Words like ``both'' and ``also'' function to draw our attention to the fact that two things are being conjoined. They are not doing any further logical work, so we do not need to represent them in SL.

Some more examples:
\begin{earg}
\item[\ex{and4}]Barbara is athletic and energetic.
\item[\ex{and5}]Barbara and Adam are both athletic.
\item[\ex{and6}]Although Barbara is energetic, she is not athletic.
\item[\ex{and7}]Barbara is athletic, but Adam is more athletic than she is.
\end{earg}

Sentence \ref{and4} is obviously a conjunction. The sentence says two things about Barbara, so in English it is permissible to refer to Barbara only once. It might be tempting to try this when translating the argument: Since $B$ means ``Barbara is athletic'', one might paraphrase the sentences as ``$B$ and energetic.'' This would be a mistake. Once we translate part of a sentence as $B$, any further structure is lost. $B$ is an atomic sentence; it is nothing more than true or false. Conversely, ``energetic'' is not a sentence; on its own it is neither true nor false. We should instead paraphrase the sentence as ``$B$ and Barbara is energetic.'' Now we need to add a sentence letter to the symbolization key. Let $E$ mean ``Barbara is energetic.'' Now the sentence can be translated as $B \eand E$.

\factoidbox{
A sentence can be symbolized as $\script{A} \eand \script{B}$ if it can be paraphrased in English as `Both \script{A}, and \script{B}.' Each of the conjuncts must be a sentence.
}

Sentence \ref{and5} says one thing about two different subjects. It says of both Barbara and Adam that they are athletic, and in English we use the word ``athletic'' only once. In translating to SL, it is important to realize that the sentence can be paraphrased as, ``Barbara is athletic, and Adam is athletic.'' This translates as $B \eand A$.

Sentence \ref{and6} is a bit more complicated. The word ``although'' sets up a contrast between the first part of the sentence and the second part. Nevertheless, the sentence says both that Barbara is energetic and that she is not athletic. In order to make each of the conjuncts an atomic sentence, we need to replace ``she'' with ``Barbara.''

So we can paraphrase sentence \ref{and6} as, ``\emph{Both} Barbara is energetic, \emph{and} Barbara is not athletic.'' The second conjunct contains a negation, so we paraphrase further: ``\emph{Both} Barbara is energetic \emph{and} \emph{it is not the case that} Barbara is athletic.'' This translates as $E \eand \enot B$.

Sentence \ref{and7} contains a similar contrastive structure. It is irrelevant for the purpose of translating to SL, so we can paraphrase the sentence as ``\emph{Both} Barbara is athletic, \emph{and} Adam is more athletic than Barbara.'' (Notice that we once again replace the pronoun ``she'' with her name.) How should we translate the second conjunct? We already have the sentence letter $A$ which is about Adam's being athletic and $B$ which is about Barbara's being athletic, but neither is about one of them being more athletic than the other. We need a new sentence letter. Let $R$ mean ``Adam is more athletic than Barbara.'' Now the sentence translates as $B \eand R$.

\factoidbox{Sentences that can be paraphrased ``\script{A}, but \script{B}'' or ``Although \script{A}, \script{B}'' are best symbolized using conjunction  \script{A} \eand \script{B}.}

It is important to keep in mind that the sentence letters $A$, $B$, and $R$ are atomic sentences. Considered as symbols of SL, they have no meaning beyond being true or false. We have used them to symbolize different English language sentences that are all about people being athletic, but this similarity is completely lost when we translate to SL. No formal language can capture all the structure of the English language, but as long as this structure is not important to the argument there is nothing lost by leaving it out.

For any sentences \script{A} and \script{B}, \script{A} \eand \script{B} is true if and only if both \script{A} and \script{B} are true. We can summarize this in the {characteristic truth table} for conjunction:
\begin{center}
\begin{tabular}{c|c|c}
\script{A} & \script{B} & \script{A} \eand \script{B}\\
\hline
T & T & T\\
T & F & F\\
F & T & F\\
F & F & F
\end{tabular}
\end{center}

Conjunction is symmetrical because we can swap the conjuncts without changing the truth value of the sentence. Regardless of what \script{A} and \script{B} are, \script{A}\eand\script{B} is logically equivalent to \script{B} \eand \script{A}.


%%%%%%%%%%%%%%%%%%%% 2.2.3 disjunction

\subsection{Disjunction}
Consider these sentences:
\begin{earg}
\item[\ex{or1}]Either Denison will play golf with me, or he will watch movies.
\item[\ex{or2}]Either Denison or Ellery will play golf with me. 
\end{earg}

For these sentences we can use this symbolization key:

\begin{ekey}
\item[D:] Denison will play golf with me.
\item[E:] Ellery will play golf with me.
\item[M:] Denison will watch movies.
\end{ekey}

\newglossaryentry{disjunction}
{
name=disjunction,
description={The symbol \eor, used to represent words and phrases that function like the English word ``or'' in its inclusive sense.}
}

\newglossaryentry{disjunct}
{
name=disjunct,
description={A sentences joined to another by a disjunction.}
}



Sentence \ref{or1} is ``Either $D$ or $M$.'' To fully symbolize this, we introduce a new symbol. The sentence becomes $D \eor M$. The \eor connective is called \textsc{\gls{disjunction}}, \label{def:disjunction} and $D$ and $M$ are called \textsc{\glspl{disjunct}}. \label{def:disjunct}

Sentence \ref{or2} is only slightly more complicated. There are two subjects, but the English sentence only gives the verb once. In translating, we can paraphrase it as ``Either Denison will play golf with me, or Ellery will play golf with me.'' Now it obviously translates as $D \eor E$.


\factoidbox{
A sentence can be symbolized as $\script{A}\eor\script{B}$ if it can be paraphrased in English as ``Either \script{A} or \script{B}.'' Each of the disjuncts must be a sentence.
}

Sometimes in English, the word ``or'' excludes the possibility that both disjuncts are true. This is called an \define{exclusive or}.  An \emph{exclusive or} is clearly intended when a restaurant menu says, ``Entrees come with either soup or salad.'' You may have soup; you may have salad; but, if you want \emph{both} soup \emph{and} salad, then you will have to pay extra.

At other times, the word ``or'' allows for the possibility that both disjuncts might be true. This is probably the case with sentence \ref{or2}, above. I might play with Denison, with Ellery, or with both Denison and Ellery. Sentence \ref{or2} merely says that I will play with \emph{at least} one of them. This is called an \define{inclusive or}.

The symbol $\eor$ represents an \emph{inclusive or}. So $D \eor E$ is true if $D$ is true, if $E$ is true, or if both $D$ and $E$ are true. It is false only if both $D$ and $E$ are false. We can summarize this with the {characteristic truth table} for disjunction:

\begin{center}
\begin{tabular}{c|c|c}
\script{A} & \script{B} & \script{A} \eor \script{B} \\
\hline
T & T & T\\
T & F & T\\
F & T & T\\
F & F & F
\end{tabular}
\end{center}

Like conjunction, disjunction is symmetrical. \script{A} \eor \script{B} is logically equivalent to \script{B} \eor \script{A}.


These sentences are somewhat more complicated:

\begin{earg}
\item[\ex{or3}] Either you will not have soup, or you will not have salad.
\item[\ex{or4}] You will have neither soup nor salad.
\item[\ex{or.xor}] You get either soup or salad, but not both.
\end{earg}

We let $S_1$ mean that you get soup and $S_2$ mean that you get salad.

Sentence \ref{or3} can be paraphrased in this way: ``Either \emph{it is not the case that} you get soup, or \emph{it is not the case that} you get salad.'' Translating this requires both disjunction and negation. It becomes $\enot S_1 \eor \enot S_2$.

Sentence \ref{or4} also requires negation. It can be paraphrased as, ``\emph{It is not the case that} either you get soup or you get salad.'' We need some way of indicating that the negation does not just negate the right or left disjunct, but rather negates the entire disjunction. In order to do this, we put parentheses around the disjunction: ``It is not the case that $(S_1 \eor S_2)$.'' This becomes simply $\enot (S_1 \eor S_2)$.

Notice that the parentheses are doing important work here. The sentence $\enot S_1 \eor S_2$ would mean ``Either you will not have soup, or you will have salad.''

Sentence \ref{or.xor} is an \emph{exclusive or}. Although $\eor$ is an inclusive or, we can symbolize an exclusive or in {SL}. We just need more than one connective to do it. We can break the sentence into two parts. The first part says that you get one or the other. We translate this as $(S_1 \eor S_2)$. The second part says that you do not get both. We can paraphrase this as ``It is not the case both that you get soup and that you get salad.'' Using both negation and conjunction, we translate this as $\enot(S_1 \eand S_2)$. Now we just need to put the two parts together. As we saw above, ``but'' can usually be translated as a conjunction. Sentence \ref{or.xor} can thus be translated as $(S_1 \eor S_2) \eand \enot(S_1 \eand S_2)$.

%%%%%%%%%%%%%%%%% 2.2.4 conditional

\subsection{Conditional}
For the following sentences, let $R$ mean ``You will cut the red wire'' and $B$ mean ``The bomb will explode.''

\begin{earg}
\item[\ex{if1}] If you cut the red wire, then the bomb will explode.
\item[\ex{if2}] The bomb will explode only if you cut the red wire.
\end{earg}

\newglossaryentry{conditional}
{
name=conditional,
description={The symbol \eif, used to represent words and phrases that function like the English phrase ``if \ldots then.''}
}

\newglossaryentry{antecedent}
{
name=antecedent,
description={The sentence to the left of a conditional..}
}


\newglossaryentry{consequent}
{
name=consequent,
description={The sentence to the right of a conditional.}
}


Sentence \ref{if1} can be translated partially as ``If $R$, then $B$.'' We will use the symbol $\eif$ to represent logical {entailment}. Sentence \ref{if1} then becomes $R\eif B$. The connective is called a \textsc{\gls{conditional}}. \label{def:conditional} The sentence on the left-hand side of the conditional ($R$ in this example) is called the \textsc{\gls{antecedent}}. \label{def:antecedent}
 The sentence on the right-hand side ($B$) is called the \textsc{\gls{consequent}}. \label{def:consequent} 

Sentence \ref{if2} is also a conditional. Since the word ``if'' appears in the second half of the sentence, it might be tempting to symbolize this in the same way as sentence \ref{if1}. That would be a mistake.

The conditional $R\eif B$ says that \emph{if} $R$ were true, \emph{then} $B$ would also be true. It does not say that you cutting the red wire is the \emph{only} way that the bomb could explode. Someone else might cut the wire, or the bomb might be on a timer. The sentence $R\eif B$ does not say anything about 
what to expect if $R$ is false. Sentence \ref{if2} is different. It says that the only conditions under which the bomb will explode involve you having cut the red wire; i.e., if the bomb explodes, then you must have cut the wire. As such, sentence \ref{if2} should be symbolized as $B \eif R$.

It is important to remember that the connective $\eif$ says only that, if the antecedent is true, then the consequent is true. It says nothing about the \emph{causal} connection between the two events. Translating sentence \ref{if2} as $B \eif R$ does not mean that the bomb exploding would somehow have caused you cutting the wire. Both sentence \ref{if1} and \ref{if2} suggest that, if you cut the red wire, you cutting the red wire would be the cause of the bomb exploding. They differ on the \emph{logical} connection. If sentence \ref{if2} were true, then an explosion would tell us---those of us safely away from the bomb---that you had cut the red wire. Without an explosion, sentence \ref{if2} tells us nothing.

\factoidbox{
The paraphrased sentence ``\script{A} only if \script{B}'' is logically equivalent to ``If \script{A}, then \script{B}.''
}

% Could discuss necessary and sufficient conditions here.

``If \script{A}, then \script{B}'' means that if \script{A} is true, then so is \script{B}. So we know that if the antecedent \script{A} is true but the consequent \script{B} is false, then the conditional ``If \script{A} then \script{B}'' is false. What is the truth value of ``If \script{A}, then \script{B}'' under other circumstances? Suppose, for instance, that the antecedent \script{A} happened to be false. ``If \script{A}, then \script{B}'' would then not tell us anything about the actual truth value of the consequent \script{B}, and it is unclear what the truth value of ``If \script{A}, then \script{B}'' would be.

In English, the truth of conditionals often depends on what \emph{would} be the case if the antecedent \emph{were true}---even if, as a matter of fact, the antecedent is false. This poses a problem for translating conditionals into SL.  Considered as sentences of SL, $R$ and $B$ in the above examples have nothing intrinsic to do with each other. In order to consider what the world would be like if $R$ were true, we would need to analyze what $R$ says about the world. Since $R$ is an atomic symbol of SL, however, there is no further structure to be analyzed. When we replace a sentence with a sentence letter, we consider it merely as some atomic sentence that might be true or false.

In order to translate conditionals into SL, we will not try to capture all the subtleties of the English language ``If$\ldots$, then$\ldots$.'' Instead, the symbol $\eif$ will be what logicians call a material conditional. This means that when \script{A} is false, the conditional \script{A} \eif \script{B} is automatically true, regardless of the truth value of \script{B}. If both \script{A} and \script{B} are true, then the conditional \script{A} \eif \script{B} is true.


In short, \script{A} \eif \script{B} is false if and only if \script{A} is true and \script{B} is false. We can summarize this with a characteristic truth table for the conditional.

\begin{center}
\begin{tabular}{c|c|c}
\script{A} & \script{B} & \script{A}\eif\script{B}\\
\hline
T & T & T\\
T & F & F\\
F & T & T\\
F & F & T
\end{tabular}
\end{center}

The conditional is asymmetrical. You cannot swap the antecedent and consequent without changing the meaning of the sentence, because \script{A} \eif \script{B} and \script{B} \eif \script{A} are not logically equivalent.

%\begin{earg}
%\item[\ex{if3}] Everytime a bell rings, an angel earns its wings.
%\item[\ex{if4}] Bombs always explode when you cut the red wire.
%\end{earg}

Not all sentences of the form ``If$\ldots$, then$\ldots$'' are conditionals. Consider this sentence:

\begin{earg}
\item[\ex{if5}] If anyone wants to see me, then I will be on the porch.
\end{earg}

When I say this, it means that I will be on the porch, regardless of whether anyone wants to see me or not---but if someone did want to see me, then they should look for me there. If we let $P$ mean ``I will be on the porch,'' then sentence \ref{if5} can be translated simply as $P$.

%%%%%%%%%%%%%%%% 6.2.5 Biconditional

\subsection{Biconditional}
Consider these sentences:
\begin{earg}
\item[\ex{iff1}] The figure on the board is a triangle only if it has exactly three sides.
\item[\ex{iff2}] The figure on the board is a triangle if it has exactly three sides.
\item[\ex{iff3}] The figure on the board is a triangle if and only if it has exactly three sides.
\end{earg}

Let $T$ mean ``The figure is a triangle'' and $S$ mean ``The figure has three sides.''

Sentence \ref{iff1}, for reasons discussed above, can be translated as $T\eif S$.

Sentence \ref{iff2} is importantly different. It can be paraphrased as ``If the figure has three sides, then it is a triangle.'' So it can be translated as $S\eif T$.

\newglossaryentry{biconditional}
{
name=biconditional,
description={The symbol \eiff, used to represent words and phrases that function like the English phrase ``if and only if.''}
}


Sentence \ref{iff3} says that $T$ is true \emph{if and only if} $S$ is true; we can infer $S$ from $T$, and we can infer $T$ from $S$. This is called a \textsc{\gls{biconditional}}, \label{def:biconditional} because it entails the two conditionals $S\eif T$ and $T \eif S$. We will use \eiff  to represent the biconditional; sentence \ref{iff3} can be translated as $S \eiff T$.

We could abide without a new symbol for the biconditional. Since sentence \ref{iff3} means ``$T \eif S$ and $S\eif T$,'' we could translate it as $(T \eif S)\eand(S\eif T)$. We would need parentheses to indicate that $(T \eif S)$ and $(S\eif T)$ are separate conjuncts; the expression $T \eif S\eand S\eif T$ would be ambiguous.

Because we could always write $(\script{A}\eif\script{B})\eand(\script{B}\eif\script{A})$ instead of $\script{A}\eiff\script{B}$, we do not strictly speaking \emph{need} to introduce a new symbol for the biconditional. Nevertheless, logical languages usually have such a symbol. SL will have one, which makes it easier to translate phrases like ``if and only if.''

\script{A} \eiff \script{B} is true if and only if \script{A} and \script{B} have the same truth value. \label{defBiconditional}This is the characteristic truth table for the biconditional:

\begin{center}
\begin{tabular}{c|c|c}
\script{A} & \script{B} & \script{A} \eiff \script{B}\\
\hline
T & T & T\\
T & F & F\\
F & T & F\\
F & F & T
\end{tabular}
\end{center}

% ******************************************
%  * 		Other Symbolization	                          *
% ******************************************

\section{Other Symbolization}
We have now introduced all of the connectives of SL. We can use them together to translate many kinds of sentences. Consider these examples of sentences that use the English-language connective ``unless'':

\begin{earg}
\item[\ex{unless1}] Unless you wear a jacket, you will catch cold. 
\item[\ex{unless2}] You will catch cold unless you wear a jacket. 
\end{earg}

Let $J$ mean ``You will wear a jacket'' and let $D$ mean ``You will catch a cold.''

We can paraphrase sentence \ref{unless1} as ``Unless $J$, $D$.'' This means that if you do not wear a jacket, then you will catch cold; with this in mind, we might translate it as $\enot J \eif D$. It also means that if you do not catch a cold, then you must have worn a jacket; with this in mind, we might translate it as $\enot D \eif J$.

Which of these is the correct translation of sentence \ref{unless1}? Both translations are correct, because the two translations are logically equivalent in SL.

Sentence \ref{unless2}, in English, is logically equivalent to sentence \ref{unless1}. It can be translated as either $\enot J \eif D$ or $\enot D \eif J$.

When symbolizing sentences like sentence \ref{unless1} and sentence \ref{unless2}, it is easy to get turned around. Since the conditional is not symmetric, it would be wrong to translate either sentence as $J \eif \enot D$. Fortunately, there are other logically equivalent expressions. Both sentences mean that you will wear a jacket or---if you do not wear a jacket---then you will catch a cold. So we can translate them as $J \eor D$. (You might worry that the ``or'' here should be an \emph{exclusive or}. However, the sentences do not exclude the possibility that you might \emph{both} wear a jacket \emph{and} catch a cold; jackets do not protect you from all the possible ways that you might catch a cold.)


\factoidbox{
If a sentence can be paraphrased as ``Unless \script{A}, \script{B},'' then it can be symbolized as $\script{A}\eor\script{B}$.
}




% ******************************************
%  * 		Recursive Syntax for  SL                    *
% ******************************************


\section{Recursive Syntax for SL} % I reworked this section to focus on the idea of recursive syntax

The previous two sections gave you a rough, informal sense of how to create sentences in SL. If I give you an English sentence like ``Grass is either green or brown,'' you should be able to write a corresponding sentence in SL: ``$A \eor B$.'' In this section we want to give a more precise definition of a sentence in SL.  When we defined sentences in English, we did so using the concept of truth: Sentences were units of language that can be true or false. In SL, it is possible to define what counts as a sentence without talking about truth. Instead, we can just talk about the structure of the sentence. This is one respect in which a formal language like SL is more precise than a natural language like English.

\newglossaryentry{syntax}
{
name=syntax,
description={The structure of a bit of language, considered without reference to truth, falsity, or meaning.}
}

\newglossaryentry{semantics}
{
name=semantics,
description={The meaning of a bit of language is its meaning, including truth and falsity.}
}

The structure of a sentence in SL considered without reference to truth or falsity is called its syntax. More generally \textsc{\gls{syntax}} \label{def:syntax} refers to the study of the properties of language that are there even when you don't consider meaning. Whether a sentence is true or false is considered part of its meaning. In this chapter, we will be giving a purely syntactical definition of a sentence in SL.  The contrasting term is \textsc{\gls{semantics}} \label{def:semantics} the study of aspects of language that relate to meaning, including truth and falsity. (The word ``semantics'' comes from the Greek word for ``mark'')

\newglossaryentry{object language}
{
name=object language,
description={A language that is constructed and studied by logicians. In this textbook, the object languages are SL and QL.}
}

\newglossaryentry{metalanguage}
{
name=metalanguage,
description={The language logicians use to talk about the object language. In this textbook, the metalanguage is English, supplemented by certain symbols like metavariables and technical terms like ``valid.''}
}

If we are going to define a sentence in SL just using syntax, we will need to carefully distinguish SL from the language that we use to talk about SL. When you create an artificial language like SL, the language that you are creating is called the \textsc{\gls{object language}}. \label{def:object_language} The language that we use to talk about the object language is called the \textsc{\gls{metalanguage}}. \label{def:metalanguage} Imagine building a house. The object language is like the house itself. It is the thing we are building. While you are building a house, you might put up scaffolding around it. The scaffolding isn't part of the the house. You just use it to build the house. The metalanguage is like the scaffolding. 

The object language in this chapter is SL. For the most part, we can build this language just by talking about it in ordinary English. However we will also have to build some special scaffolding that is not a part of SL, but will help us build SL. Our metalanguage will thus be ordinary English plus this scaffolding.

\newglossaryentry{metavariables}
{
name=metavariables,
description={A variable in the metalanguage that can represent any sentence in the object language.}
}



%rob: Paragraph on metavariables added.
An important part of the scaffolding are the \textsc{\gls{metavariables}} \label{def:metavariables} These are the fancy script letters we have been using in the characteristic truth tables for the connectives: \script{A}, \script{B}, \script{C}, etc. These are letters that can refer to any sentence in SL. They can represent sentences like $P$ or $Q$, or they can represent longer sentences, like $(((A \eor B) \eand G) \eif (P \eiff Q))$. Just as the sentence letters $A$, $B$, etc. are variables that range over any English sentence, the metavariables \script{A}, \script{B}, etc. are variables that range over any sentence in SL, including the sentence letters $A$, $B$, etc. 

As we said, in this chapter we will give a syntactic definition for ``sentence of SL.'' The definition itself will be given in mathematical English, the metalanguage.

\begin{center}
\begin{tabular}{|c|c|}
\hline
sentence letters & $A,B,C,\ldots,Z$\\
with subscripts, as needed & $A_1, B_1,Z_1,A_2,A_{25},J_{375},\ldots$\\
\hline
connectives & \enot,\eand,\eor,\eif,\eiff\\
\hline
parentheses&( , )\\
\hline
\end{tabular}
\end{center}

Most random combinations of these symbols will not count as sentences in SL. Any random connection of these symbols will just be called a ``string'' or ``expression'' Random strings only become meaningful sentences when the are structured according to the rules of syntax. We saw from the earlier two sections that individual sentence letters,  like $A$ and $G_{13}$ counted as sentences. We also saw that we can put these sentences together using connectives so that  $\enot A$ and $\enot G_{13}$ is a sentence.  The problem is, we can't simply list all the different sentences we can put together this way, because there are infinitely many of them. Instead, we will define a sentence in SL by specifying the process by which they are constructed.

Consider negation: Given any sentence \script{A} of SL, $\enot\script{A}$ is a sentence of SL. It is important here that \script{A} is not the sentence letter $A$. Rather, it is a metavariable: part of the metalanguage, not the object language. Since \script{A} is not a symbol of SL, $\enot\script{A}$ is not an expression of SL. Instead, it is an expression of the metalanguage that allows us to talk about infinitely many expressions of SL: all of the expressions that start with the negation symbol. 


\newglossaryentry{sentence of SL}
{
name=sentence of SL,
description={A string of symbols in SL that can be built up using according to the recursive rules given on page }
}



We can say similar things for each of the other connectives. For instance, if \script{A} and \script{B} are sentences of SL, then $(\script{A}\eand\script{B})$ is a sentence of SL. Providing clauses like this for all of the connectives, we arrive at the following formal definition for a \textsc{\gls{sentence of SL}}: \label{def:sentence_of_SL}

\begin{enumerate}
\item Every atomic sentence is a sentence.
\item If \script{A} is a sentence, then $\enot\script{A}$ is a sentence of SL.
\item If \script{A} and \script{B} are sentences, then $(\script{A}\eand\script{B})$ is a sentence.
\item If \script{A} and \script{B} are sentences, then $(\script{A}\eor\script{B})$ is a sentence.
\item If \script{A} and \script{B} are sentences, then $(\script{A}\eif\script{B})$ is a sentence.
\item If \script{A} and \script{B} are sentences, then $(\script{A}\eiff\script{B})$ is a sentence.
\item All and only sentences of SL can be generated by applications of these rules.
\end{enumerate}

We can apply this definition to see whether an arbitrary string is a sentence. Suppose we want to know whether or not $\enot \enot \enot D$ is a sentence of SL. Looking at the second clause of the definition, we know that $\enot \enot \enot D$ is a sentence \emph{if} $\enot \enot D$ is a sentence. So now we need to ask whether or not $\enot \enot D$ is a sentence. Again looking at the second clause of the definition, $\enot \enot D$ is a sentence \emph{if} $\enot D$ is. Again, $\enot D$ is a sentence \emph{if} $D$ is a sentence. Now $D$ is a sentence letter, an atomic sentence of SL, so we know that $D$ is a sentence by the first clause of the definition. So for a compound formula like $\enot \enot \enot D$, we must apply the definition repeatedly. Eventually we arrive at the atomic sentences from which the sentence is built up.

\newglossaryentry{recursive definition}
{
name=recursive definition,
description={A definition that defines a term by identifying base class and rules for extending that class.}
}

Definitions like this are called recursive. \textsc{\Glspl{recursive definition}} \label{def:recursive_definition} begin with some specifiable base elements and define ways to indefinitely compound the base elements. Just as the recursive definition allows complex sentences to be built up from simple parts, you can use it to decompose sentences into their simpler parts. To determine whether or not something meets the definition, you may have to refer back to the definition many times.

\newglossaryentry{scope}
{
name=scope,
description={The sentences that are joined by a connective. These are the sentences the connective was applied to when the sentence was assembled using a recursive definition.}
}

When you use a connective to build a longer sentence from shorter ones, the shorter sentences are said to be in the \textsc{\gls{scope}} \label{def:scope} of the connective. So in the sentence $(A \eand B) \eif C$, the scope of the connective $\eif$ includes $(A \eand B)$ and C. In the sentence $\enot(A \eand B)$ the scope of the $\enot$ is $(A \eand B)$. On the other hand, in the sentence $\enot A \eand B$ the scope of the $\enot$ is just A.

\newglossaryentry{main connective}
{
name=main connective,
description={The last connective that you add when you assemble a sentence using the recursive definition.}
}

The last connective that you add when you assemble a sentence using the recursive definition is the \textsc{\gls{main connective}} \label{def:main_connective} of that sentence. For example: The main logical operator of $\enot (E \eor (F \eif G))$ is negation, \enot. The main logical operator of $(\enot E \eor (F \eif G))$ is disjunction, \eor. The main connective of any sentence will have all the rest of the sentence in its scope.

%The recursive structure of sentences in SL will be important when we consider the circumstances under which a particular sentence would be true or false. The sentence $\enot \enot \enot D$ is true if and only if the sentence $\enot \enot D$ is false, and so on through the structure of the sentence until we arrive at the atomic components: $\enot \enot \enot D$ is true if and only if the atomic sentence $D$ is false. We will return to this point in the next chapter.
%restore when you restore the recursive part of chap. 3.

\subsection{Notational conventions}
\label{SLconventions}
A sentence like $(Q \eand R)$ must be surrounded by parentheses, because we might apply the definition again to use this as part of a more complicated sentence. If we negate $(Q \eand R)$, we get $\enot(Q \eand R)$. If we just had $Q \eand R$ without the parentheses and put a negation in front of it, we would have $\enot Q \eand R$. It is most natural to read this as meaning the same thing as $(\enot Q \eand R)$, something very different than $\enot(Q\eand R)$. The sentence $\enot(Q \eand R)$ means that it is not the case that both $Q$ and $R$ are true; $Q$ might be false or $R$ might be false, but the sentence does not tell us which. The sentence $(\enot Q \eand R)$ means specifically that $Q$ is false and that $R$ is true. As such, parentheses are crucial to the meaning of the sentence.

So, strictly speaking, $Q \eand R$ without parentheses is \emph{not} a sentence of SL. When using SL, however, we will often be able to relax the precise definition so as to make things easier for ourselves. We will do this in several ways.

First,  we understand that $Q \eand R$ means the same thing as $(Q \eand R)$. As a matter of convention, we can leave off parentheses that occur \emph{around the entire sentence}.

Second, it can sometimes be confusing to look at long sentences with many nested pairs of parentheses. We adopt the convention of using square brackets [ and ] in place of parentheses. There is no logical difference between $(P\eor Q)$ and $[P\eor Q]$, for example. The unwieldy sentence
$$(((H \eif I) \eor (I \eif H)) \eand (J \eor K))$$
could be written in this way:
$$\bigl[(H \eif I) \eor (I \eif H)\bigr] \eand (J \eor K)$$


Third, we will sometimes want to translate the conjunction of three or more sentences. For the sentence ``Alice, Bob, and Candice all went to the party,'' suppose we let $A$ mean ``Alice went,'' $B$ mean ``Bob went,'' and $C$ mean ``Candice went.'' The definition only allows us to form a conjunction out of two sentences, so we can translate it as $(A \eand B) \eand C$ or as $A \eand (B \eand C)$. There is no reason to distinguish between these, since the two translations are logically equivalent. There is no logical difference between the first, in which $(A \eand B)$ is conjoined with $C$, and the second, in which $A$ is conjoined with $(B \eand C)$.  So we might as well just write $A \eand B \eand C$. As a matter of convention, we can leave out parentheses when we conjoin three or more sentences.

Fourth, a similar situation arises with multiple disjunctions. ``Either Alice, Bob, or Candice went to the party'' can be translated as $(A \eor B) \eor C$ or as $A \eor (B \eor C)$. Since these two translations are logically equivalent, we may write $A \eor B \eor C$.

These latter two conventions only apply to multiple conjunctions or multiple  disjunctions. If a series of connectives includes both disjunctions and conjunctions, then the parentheses are essential; as with $(A \eand B) \eor C$ and $A \eand (B \eor C)$. The parentheses are also required if there is a series of conditionals or biconditionals; as with $(A \eif B) \eif C$ and $A \eiff (B \eiff C)$.

We have adopted these four rules as notational conventions, not as changes to the definition of a sentence. Strictly speaking, $A \eor B \eor C$ is still not a sentence. Instead, it is a kind of shorthand. We write it for the sake of convenience, but we really mean the sentence $(A \eor (B \eor C))$.

If we had given a different definition for a sentence, then these could count as sentences. We might have written rule 3 in this way: ``If \script{A}, \script{B}, $\ldots$ \script{Z} are sentences, then $(\script{A}\eand\script{B}\eand\ldots\eand\script{Z})$, is a sentence .'' This would make it easier to translate some English sentences, but would have the cost of making our formal language more complicated. We would have to keep the complex definition in mind when we develop truth tables and a proof system. We want a logical language that is expressively simple and allows us to translate easily from English, but we also want a formally simple language. Adopting notational conventions is a compromise between these two desires.


\practiceproblems
\setlength{\parindent}{0em}

\problempart Using the symbolization key given, translate each English-language sentence into SL.
\label{pr.monkeysuits}
\begin{ekey}
\item[M:] Those creatures are men in suits. 
\item[C:] Those creatures are chimpanzees. 
\item[G:] Those creatures are gorillas.
\end{ekey}
\begin{earg}
\item Those creatures are not men in suits. % {\color{red}$\enot M$} \vspace{1ex}
\item Those creatures are men in suits, or they are not. % {\color{red}$M \eor \enot M$} \vspace{1ex}
\item Those creatures are either gorillas or chimpanzees. % {\color{red}$G \eor C$} \vspace{1ex}
\item Those creatures are not gorillas, but they are not chimpanzees either. % {\color{red}$\enot G \eand \enot C$} \vspace{1ex}
\item Those creatures cannot be both gorillas and men in suits. % {\color{red}$\enot(G \eand M)$} \vspace{1ex}
\item If those creatures are not gorillas, then they are men in suits % {\color{red} $\enot G \eif M$} \vspace{1ex}
\item Those creatures are men in suits only if they are not gorillas. % {\color{red} $M \eif \enot G$} \vspace{1ex}
\item Those creatures are chimpanzees if and only if they are not gorillas. % {\color{red} $C \eiff G$} \vspace{1ex}
\item Those creatures are neither gorillas nor chimpanzees. % {\color{red} $~(G \eor C).$ See p.34, sentence 19, and p. 156} \vspace{1ex}
\item Unless those creatures are men in suits, they are either chimpanzees or they are gorillas. % {\color{red}$M \eor (C \eor G)$} \vspace{1ex}
\end{earg}

%If					X
%only if				X
%if and only if		X
%but				X
%unless				X
%not both      		X
%neither nor			X

%added and changed problems to get a better distribution of kinds of problems. 

\problempart Using the symbolization key given, translate each English-language sentence into SL.
\begin{ekey}
\item[A:] Mister Ace was murdered.
\item[B:] The butler did it.
\item[C:] The cook did it.
\item[D:] The Duchess is lying.
\item[E:] Mister Edge was murdered.
\item[F:] The murder weapon was a frying pan.
\end{ekey}
\begin{earg}
\item Either Mister Ace or Mister Edge was murdered. % {\color{red} $A \eor E$}  \vspace{1ex}
\item If Mister Ace was murdered, then the cook did it. % {\color{red} $A \eif C$} \vspace{1ex}
\item If Mister Edge was murdered, then the cook did not do it. % {\color{red} $E \eif \enot C} \vspace{1ex}
\item Either the butler did it, or the Duchess is lying. % {\color{red} $B \eor D$} \vspace{1ex}
\item The cook did it only if the Duchess is lying. % {\color{red} $C \eif D$} \vspace{1ex}
\item If the murder weapon was a frying pan, then the culprit must have been the cook. % {\color{red} $F \eif C$} \vspace{1ex}
\item If the murder weapon was not a frying pan, then the culprit was neither the cook nor the butler. % {\color{red} $\enot F \eif \enot(C \or B) \vspace{1ex}
\item Mister Ace was murdered if and only if Mister Edge was not murdered. % {\color{red} $A \eiff \enot E$} \vspace{1ex}
\item The Duchess is lying, unless it was Mister Edge who was murdered. % {\color{red} $D \eor A$} \vspace{1ex}
\item Mister Ace was murdered, but not with a frying pan. % {\color{red} $A \eand \enot F$} \vspace{1ex}
\item The butler and the cook did not both do it. % {\color{red} $\enot(B \enad C)$} \vspace{1ex}
\item Of course the Duchess is lying! % {\color{red}$D$} \vspace{1ex}
\end{earg}

%If  			x
%only if             x
%if and only if   x
%but			x
%unless			x
%not both		x
%neither nor		x

%changed problems to get a better distribution of kinds of problems. 


\problempart Using the symbolization key given, translate each English-language sentence into SL.
\label{pr.avacareer}
\begin{ekey}
\item[E$_1$:] Ava is an electrician.
\item[E$_2$:] Harrison is an electrician.
\item[F$_1$:] Ava is a firefighter.
\item[F$_2$:] Harrison is a firefighter.
\item[S$_1$:] Ava is satisfied with her career.
\item[S$_2$:] Harrison is satisfied with his career.
\end{ekey}
\begin{earg}
\item Ava and Harrison are both electricians. %{\color{red} $E_1 \eand E_2$} \vspace{1ex}
\item If Ava is a firefighter, then she is satisfied with her career. %{\color{red} $F_1 \eif S_1$}  \vspace{1ex}
\item Ava is a firefighter, unless she is an electrician. %{\color{red} $F_1 \eor E_1$  \vspace{1ex}
\item Harrison is an unsatisfied electrician. %{\color{red} $E_2 \eand \enot S_2$}  \vspace{1ex}
\item Neither Ava nor Harrison is an electrician. %{\color{red} $\enot(E_1 \eor E_2)$}  \vspace{1ex}
\item Both Ava and Harrison are electricians, but neither of them find it satisfying. %{\color{red} $(E_1 \eand E_2) \eand \enot (S_1 \eor S_2)$} \vspace{1ex}
\item Harrison is satisfied only if he is a firefighter. %{\color{red} $S_2 \eif F_2$} \vspace{1ex}
\item If Ava is not an electrician, then neither is Harrison, but if she is, then he is too. %{\color{red} $(\enot E_1 \eif \enot E_2) \eand (E_1 \eif E_2)$} \vspace{1ex}
\item Ava is satisfied with her career if and only if Harrison is not satisfied with his. %{\color{red} $S_1 \eiff \enot S_2$} \vspace{1ex}
\item If Harrison is both an electrician and a firefighter, then he must be satisfied with his work. %{\color{red} $(E_2 \eand F_2) \eif S_2$} \vspace{1ex}
\item It cannot be that Harrison is both an electrician and a firefighter. %{\color{red} $\enot (E_2 \eand F_2)$} \vspace{1ex}
\item Harrison and Ava are both firefighters if and only if neither of them is an electrician. %{\color{red} $(F_1 \eand F_2) \eiff \enot (E_1 \eor E_2)$} \vspace{1ex}
\end{earg}

%If					x	
%only if				x
%if and only if		x
%but				x
%unless				x
%not both			x
%neither nor			x

\problempart Using the symbolization key given, translate each English-language sentence into SL.
\label{pr.jazzinstruments}
\begin{ekey}
\item[J$_1$:] John Coltrane played tenor sax.
\item[J$_2$:] John Coltrane played soprano sax.
\item[J$_3$:] John Coltrane played tuba
\item[M$_1$:] Miles Davis played trumpet
\item[M$_2$:]Miles Davis played tuba
\end{ekey}

\begin{earg}
\item John Coltrane played tenor and soprano sax. %{\color{red} $J_1 \eand J_2$} \vspace{1ex}
\item Neither Miles Davis nor John Coltrane played tuba. %{\color{red} $\enot(M_2 \eor J_3)$ or $\enot M_2 \eand \enot J_3$} \vspace{1ex}
\item John Coltrane did not play both tenor sax and tuba.  %{\color{red} $\enot(J_1 \eand J_3)$ or $\enot J_1 \eor \enotJ_3$} \vspace{1ex}
\item John Coltrane did not play tenor sax unless he also played soprano sax. %{\color{red} $\enot J_1 \eor J_2$} \vspace{1ex}
\item John Coltrane did not play tuba, but Miles Davis did. %{\color{red} $\enotJ_3 \eand M_2$} \vspace{1ex}
\item Miles Davis played trumpet only if he also played tuba. %{\color{red} $M_1 \eiff M_2$} \vspace{1ex}
\item If Miles Davis played trumpet, then John Coltrane played at least one of these three instruments: tenor sax, soprano sax, or tuba. %{\color{red} $M_1 \eif (J_1 \eor (J_2 \eor J_3))&} \vspace{1ex}
\item If John Coltrane played tuba then Miles Davis played neither trumpet nor tuba. %{\color{red} $J_3 \eif \enot(M_1 \eor M_2)$ or $J_3 \eif (\enot M_1 \eand \enot M_2)$  } \vspace{1ex}
\item Miles Davis and John Coltrane both played tuba if and only if Coltrane did not play tenor sax and Miles Davis did not play trumpet. %{\color{red} $(J_3 \eand M_2) \eiff \enotJ_1 & \enot M_1)$ or $(J_3 \eand M_2) \eiff \enot (J_1 \eor M_1)$} \vspace{1ex}
\end{earg}
%If					x					
%only if				x		
%if and only if		x
%but				x
%unless				x
%not both			x
%neither nor			x


\problempart
\label{pr.spies}
Give a symbolization key and symbolize the following sentences in SL. \\
%\begin{ekey}
%\item[A:] Alice is a spy
%\item [B:] Bob is a spy
%\item [C:] The code has been broken
%\item [D:] The German embassy is in an uproar
%\end{ekey}
\begin{earg}
\item Alice and Bob are both spies. % {\color{red}$A \eand B$ \vspace{1ex}}
\item If either Alice or Bob is a spy, then the code has been broken. %{\color{red} $(A \eor B) \eif C$ \vspace{1ex}}
\item If neither Alice nor Bob is a spy, then the code remains unbroken.%{\color{red}$\enot(A \eor B) \eif \enot C$ \vspace{1ex}}
\item The German embassy will be in an uproar, unless someone has broken the code.% {\color{red}$D \eor C$ \vspace{1ex}}
\item Either the code has been broken or it has not, but the German embassy will be in an uproar regardless. %{\color{red}$(C \eor \enot C) \eand D$ \vspace{1ex}}
\item Either Alice or Bob is a spy, but not both. %{\color{red}$(A \eor B) \eand \enot (A \eand B)$}
\end{earg}

%If
%only if
%if and only if
%but
%unless
%not both
%neither nor

\problempart Give a symbolization key and symbolize the following sentences in SL.
%\begin{ekey}
%\item[A:] Gregor plays first base
%\item[B:] The team will lose
%\item[C:] There is a miracle
%\item[D:] Gregor's mom will bake cookies.
%\end{ekey}

\begin{earg}
\item If Gregor plays first base, then the team will lose. %{\color{red} $A \eif B$ \vspace{1ex}}
\item The team will lose unless there is a miracle. %{\color{red}$B \eor C$ \vspace{1ex}}
\item The team will either lose or it won't, but Gregor will play first base regardless. % {\color{red}$(B \eor \enot B) \eand A$ \vspace{1ex}}
\item Gregor's mom will bake cookies if and only if Gregor plays first base.% {\color{red}$C \eiff A$ \vspace{1ex}}
\item If there is a miracle, then Gregor's mom will not bake cookies. %{\color{red} $C \eif \enot D$}
\end{earg}


\problempart
For each argument, write a symbolization key and translate the argument as well as possible into SL.
\begin{earg}
\item If Dorothy plays the piano in the morning, then Roger wakes up cranky. Dorothy plays piano in the morning unless she is distracted. So if Roger does not wake up cranky, then Dorothy must be distracted.
%{\color{red}
%\begin{ekey}
%\item[A:] Dorothy plays the piano in the morning
%\item[B:] Roger wakes up cranky
%\item[C:] Dorothy is distracted
%\end{ekey}

%begin{\earg}
%\item[1.] $A \ief B$
%\item[2.] $A \eor C$
%\item[$$\therefore$$] $\enot B \eif C$
%}
\item It will either rain or snow on Tuesday. If it rains, Neville will be sad. If it snows, Neville will be cold. Therefore, Neville will either be sad or cold on Tuesday.

%{\color{red}
%\begin{ekey}
%\item[A:]  It will rain on Tuesday
%\item[B:]  It will snow on Tuesday
%\item[C:]  Neville will be sad
%\item[D:]  Neville will be cold
%\end{ekey}

%\begin{earg}
%\item[1.]  $A \eor B$
%\item[2.]  $A \eif C$
%\item[3.] $B \eif D$
%\item[$$\therefore$$]  $C \eor D$
%\end{earg}
%}

\item If Zoog remembered to do his chores, then things are clean but not neat. If he forgot, then things are neat but not clean. Therefore, things are either neat or clean---but not both.
\end{earg}

%{\color{red}
%\begin{ekey}
%\item[A:] Zoog remembered to do his chores. 
%\item[B:] Things are clean 
%\item[C:] Things are neat %\end{ekey}

%\begin{earg}
%\item[1.]  $A \eif (B \eand \enot C)$
%\item[2.]  $\enot A \eif (\enot B \eand C)$
%\item[$\therefore$]  $(B \eor C) \eand \enot (B \eand C)$ 
%\end{earg}
%}

\problempart
For each argument, write a symbolization key and translate the argument as well as possible into SL. The part of the passage in italics is there to provide context for the argument, and doesn't need to be symbolized.
\begin{earg}
\item It is going to rain soon. I know because my leg is hurting, and my leg hurts if its going to rain. 

%{\color{red}
%\begin{ekey}
%\item[A:]  
%\item[B:]  
%\item[C:]  %\end{ekey}

%begin{\earg}
%\item[1.]  
%\item[2.]  
%\item[$\therefore$]  
%}

\item  \emph{Spider-man tries to figure out the bad guys plan.} If Doctor Octopus gets the uranium, he will blackmail the city. I am certain of this because if Doctor Octopus gets the uranium, he can make a dirty bomb, and if he can make a dirty bomb, he will blackmail the city.

%{\color{red}
%\begin{ekey}
%\item[A:]  
%\item[B:]  
%\item[C:]  %\end{ekey}

%begin{\earg}
%\item[1.]  
%\item[2.]  
%\item[$\therefore$]  
%}

\item \emph{A westerner tries to predict the policies of the Chinese government.} If the Chinese government cannot solve the water shortages in Beijing, they will have to move the capital. They dont want to move the capital. Therefore they must solve the water shortage. But the only way to solve the water shortage is to divert almost all the water from the Yangzi river northward. Therefore the Chinese government will go with the project to divert water from the south to the north.       



%{\color{red}
%\begin{ekey}
%\item[A:]  
%\item[B:]  
%\item[C:]  %\end{ekey}

%begin{\earg}
%\item[1.]  
%\item[2.]  
%\item[$\therefore$]  
%}

\end{earg}




\problempart
\begin{earg}
\item Are there any sentences of SL that contain no sentence letters? Why or why not? % \vspace{1ex} \\ {\color{red} No, because the rules for creating sentences begin with sentence letters and then apply connectives and more sentence letters. There is no way to remove the sentence letters that you start with.} \vspace{2ex} 
\item In the chapter, we symbolized an \emph{exclusive or} using \eor, \eand, and \enot. How could you translate an \emph{exclusive or} using only two connectives? Is there any way to translate an \emph{exclusive or} using only one kind of connective? %\vspace{1ex} \\ {\color{red} The exclusive or (sometimes written xor) is true whenever the two sides of it have opposite truth values. This is the reverse of what the biconditional does. Thus you can represent the xor like this: \enot(A \eiff B). You can't get rid of any more connectives, though. If you had a single connective in the sentence, it would have to be equivalent on its own to the exclusive or, and none of our connectives work like that. Some systems do introduce a separate symbol for the exclusive or, often a plus sign: +.}
\end{earg}




%\solutions
%\problempart
%\label{pr.wiffSL}
%For each of the following: (a) Is it a wff of SL? (b) Is it a sentence of SL, allowing for notational conventions?
%\begin{earg}
%\item $(A)$
%\item $J_{374} \eor \enot J_{374}$
%\item $\enot \enot \enot \enot F$
%\item $\enot \eand S$
%\item $(G \eand \enot G)$
%\item $\script{A} \eif \script{A}$
%\item $(A \eif (A \eand \enot F)) \eor (D \eiff E)$
%\item $[(Z \eiff S) \eif W] \eand [J \eor X]$
%\item $(F \eiff \enot D \eif J) \eor (C \eand D)$
%\end{earg}


%%%%    Key term list
\section*{Key Terms}
\begin{multicols}{2}
\begin{sortedlist}
\sortitem{Sentence letter}{} 	
\sortitem{Symbolization key}{} 	
\sortitem{Atomic sentence}{}
\sortitem{Sentential connective}{}
\sortitem{Negation}{}
\sortitem{Conjunction}{}
\sortitem{Conjunct}{}
\sortitem{Disjunction}{}
\sortitem{Disjunct}{}
\sortitem{Conditional}{}
\sortitem{Antecedent}{}
\sortitem{Consequent}{}
\sortitem{Biconditional}{}
\sortitem{Syntax}{}
\sortitem{Semantics}{}
\sortitem{Object language}{}
\sortitem{Metalanguage}{}
\sortitem{Metavariables}{}
\sortitem{Sentence of SL}{}
\sortitem{Main connective}{}
\sortitem{Recursive definition}{}
\sortitem{Scope}{}
\end{sortedlist}
\end{multicols}





